Elastic Scattering in General Relativistic Ray Tracing for Neutrinos

### 2016-10-26 paper.tex
Expand scattering: Legendre expansion of kernel

\subsection{Boltzmann Equation}
\label{ssec:boltzmann}
Neutrino radiation obeys the relativistic Boltzmann Equation
\todo{point out limit of QKEs}
\begin{equation}
  \label{eqn:boltzmann}
  \frac{d}{d\lambda}f(x^\alpha;\varepsilon,\upsilon_1,\upsilon_2) = C[f],
\end{equation}
where $\frac{d}{d\lambda}$ denotes a derivative with respect to the affine
parameter defining the neutrino momentum (Eqn.~\ref{eqn:geodesic_x})
and $C[f]$ is the source term arising from interactions with the medium.
The source term varies over phase space $(x^\alpha,p_\beta)$,
and depends locally on the distribution function $f$
and nonlocally on the distribution function of this neutrino and its
antiparticle at different momenta, $f'$ and $\bar{f}'$.
For clarity we symbolize all of these dependencies with the shorthand $C[f]$.
The various neutrino interactions contributing to $C[f]$ are detailed in
App.~\ref{sec:source_terms}.

@@

\subsection{Moments of the Distribution Function}
\label{ssec:moments}
\todo{code these up}
We may take angular moments of the distribution function:
\begin{align}
  \label{eqn:J}
  J(\varepsilon) &=
  \varepsilon^3 \oint d\Omega' f(\varepsilon, \ell'_\beta) \\
  \label{eqn:Ha}
  H^\mu(\varepsilon) &=
  \varepsilon^3 \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell^\mu \\
  \label{eqn:Sab}
  S^{\mu\gamma}(\varepsilon) &=
  \varepsilon^3 \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell^\mu \ell^\gamma,
\end{align}
representing specific energy density, specific energy flux, and
specific momentum flux, respectively.
Here ``specific'' refers to the quantity being integrable over neutrino energy.
\todo{relate moments to other frames?}
Integrals~\ref{eqn:J}--\ref{eqn:Sab} are performed over a solid angle in
momentum space while holding $\varepsilon$ constant:
$d\Omega \equiv d(\cos\upsilon_1)\,d\upsilon_2$.
\todo{consistent with Eqn.~\ref{eqn:def_direction}?}

@@

We follow
\cite{brue1985-core_collapse} and \cite[Sec.~4]{shib2011-truncated_moment}
by separating these processes into four categories:
\begin{equation}
  \label{eqn:four_sources}
  C[f] \equiv C_{\rm AE} + C_{\rm SE} + C_{\rm SI} + C_{\rm TP},
\end{equation}
representing charged-current absorption and emission, elastic scattering,
inelastic scattering, and thermal pair annihilation and production.
We present formulae for all of these processes, though in this work, we only
treat absorption/emission and elastic scattering.
Each term
is computed by summing the weak interaction rates of the processes from the
relevant category given in Tab.~\ref{tab:neutrino_processes}.
These rates are computed most easily as traditional emissivities and opacities
in the fluid frame of reference;
but since we use them in the form of an invariant emissivity and opacity
(defined e.g. in \ref{eqn:inv_emissivity} and \ref{eqn:inv_opacity},
and obeying Eqn.~\ref{eqn:boltzmann_linear})
they take the same numerical value in any frame of reference
and are completely independent of our choice of fiducial observer $u^\alpha$.
\todo{confirm}

@@

\subsection{Elastic scattering}
\label{ssec:sources_es}
Neutrino scattering on particles of mass much greater than $\varepsilon$
is essentially iso-energetic
(e.g. \cite[Eqn.~A8]{brue1985-core_collapse}).
\begin{equation}
  \label{eqn:sources_es}
  C_{{\rm SE},i}
  = \mathscr{K}_{{\rm SE},i} \Phi_{{\rm SE},i} - \mathscr{K}_{{\rm SE},i} f,
\end{equation}
with
\begin{align}
  \label{eqn:es_factor}
  \Phi_{{\rm SE},i} &\equiv \frac{\varepsilon^3}{\mathscr{K}_{{\rm SE},i}}
  \oint d\Omega' R(\ell_\alpha,\ell'_\beta)f(\ell'_\beta) \\
  \label{eqn:es_opacity}
  \mathscr{K}_{{\rm SE},i} &\equiv \varepsilon^3
  \oint d\Omega' R(\ell_\alpha,\ell'_\beta),
\end{align}
where $R(\ell_\alpha,\ell'_\beta)$ is the scattering kernel from direction
$\ell'_\beta$ to direction $\ell_\alpha$
having dimension $[R]={\rm energy}^{-1}$.

\subsection{Inelastic scattering}
\label{ssec:sources_is}
(e.g. \cite[Eqn.~A6]{brue1985-core_collapse}).

>>>>

\subsection{Boltzmann Equation}
\label{ssec:boltzmann}
Neutrino radiation obeys the relativistic Boltzmann Equation
\todo{point out limit of QKEs}
\begin{equation}
  \label{eqn:boltzmann}
  \frac{d}{d\lambda}f(x^\alpha;\varepsilon,\upsilon_1,\upsilon_2) = C[f],
\end{equation}
where $\frac{d}{d\lambda}$ denotes a derivative with respect to the affine
parameter defining the neutrino momentum (Eqn.~\ref{eqn:geodesic_x})
and $C[f]$ is the source term arising from interactions with the medium,
having dimension ${\rm energy}\,{\rm length}^{-1}$.
The source term varies over phase space $(x^\alpha,p_\beta)$,
and depends locally on the distribution function $f$
and nonlocally on the distribution function of this neutrino and its
antiparticle at different momenta, $f'$ and $\bar{f}'$.
For clarity we symbolize all of these dependencies with the shorthand $C[f]$.
The various neutrino interactions contributing to $C[f]$ are detailed in
App.~\ref{sec:source_terms}.

@@

\subsection{Moments of the Distribution Function}
\label{ssec:moments}
\todo{code these up}
We may take angular moments of the distribution function:
\begin{align}
  \label{eqn:J}
  J(\varepsilon) &=
  \frac{\varepsilon^3}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \\
  \label{eqn:Ha}
  H^\mu(\varepsilon) &=
  \frac{\varepsilon^3}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell^\mu \\
  \label{eqn:Sab}
  S^{\mu\gamma}(\varepsilon) &=
  \frac{\varepsilon^3}{(2\pi)^3} \oint d\Omega' f(\varepsilon, \ell'_\beta) \ell^\mu \ell^\gamma,
\end{align}
representing specific energy density, specific energy flux, and
specific momentum flux, respectively.
Here ``specific'' refers to the quantity being integrable over neutrino energy.
\todo{relate moments to other frames?}
Integrals~\ref{eqn:J}--\ref{eqn:Sab} are performed over a solid angle in
momentum space while holding $\varepsilon$ constant:
$d\Omega \equiv d(\cos\upsilon_1)\,d\upsilon_2$.
\todo{consistent with Eqn.~\ref{eqn:def_direction}?}

The energy-integrated moments are
\begin{equation}
  \label{eqn:J_H_S_eps_integrated}
  \big\{ J,H^\mu,S^{\mu\nu} \big\} = \int_0^\infty d\varepsilon \,
  \big\{ J(\varepsilon),H^\mu(\varepsilon),S^{\mu\nu}(\varepsilon) \big\},
\end{equation}
having dimension ${\rm energy}\,{\rm length}^{-3}$.

@@

We follow
\cite{brue1985-core_collapse} and \cite[Sec.~4]{shib2011-truncated_moment}
by separating these processes into four categories:
\begin{equation}
  \label{eqn:four_sources}
  C[f] \equiv C_{\rm AE} + C_{\rm SE} + C_{\rm SI} + C_{\rm PP},
\end{equation}
representing charged-current absorption and emission, elastic scattering,
inelastic scattering, and the thermal pair processes of annihilation and
production.
We present formulae for all of these processes, though in this work, we only
treat absorption/emission and elastic scattering.
Each term
is computed by summing the weak interaction rates of the processes from the
relevant category given in Tab.~\ref{tab:neutrino_processes}.
These rates are computed most easily as traditional emissivities and opacities
in the fluid frame of reference;
but since we use them in the form of an invariant emissivity and opacity
(defined e.g. in \ref{eqn:inv_emissivity} and \ref{eqn:inv_opacity},
and obeying Eqn.~\ref{eqn:boltzmann_linear})
they take the same numerical value in any frame of reference
and are completely independent of our choice of fiducial observer $u^\alpha$.
\todo{confirm}

@@

\subsection{Elastic scattering}
\label{ssec:sources_se}
Neutrino scattering on particles of mass much greater than $\varepsilon$
is essentially iso-energetic. Following 
\cite[Eqn.~A8]{brue1985-core_collapse} or
\cite[Eqn.~4.20]{shib2011-truncated_moment} the collision term for the
$i$-th process takes the form
\begin{equation}
  \label{eqn:integral_se}
  C_{{\rm SE},i}
  = \frac{\varepsilon^3}{(2\pi)^3}
  \oint d\Omega' R_{{\rm SE},i}(\omega')
  \big(f(\ell'_\beta) - f(\ell_\alpha)\big)
\end{equation}
where $R(\omega')$ is the scattering kernel from direction $\ell'_\beta$
to direction $\ell_\alpha$ having dimension $[R]={\rm energy}^{-1}$,
and the cosine of the scattering angle is
$\omega' \equiv 1-\ell'_\alpha \ell^\alpha$.
We write this in the form
\begin{equation}
  \label{eqn:sources_se}
  C_{{\rm SE},i}
  = \mathscr{K}_{{\rm SE},i} \Phi_{{\rm SE},i} - \mathscr{K}_{{\rm SE},i} f,
\end{equation}
with
\begin{align}
  \label{eqn:se_factor}
  \Phi_{{\rm SE},i} &\equiv
  \frac{1}{\mathscr{K}_{{\rm SE},i}}
  \frac{\varepsilon^3}{(2\pi)^3}
  \oint d\Omega' R_{{\rm SE},i}(\omega')f(\ell'_\beta) \\
  \label{eqn:se_opacity}
  \mathscr{K}_{{\rm SE},i} &\equiv
  \frac{\varepsilon^3}{(2\pi)^3}
  \oint d\Omega' R_{{\rm SE},i}(\omega').
\end{align}

It is customary to approximate the scattering kernel to linear order in
$\omega$ (as in~\cite[Eqn.~4.21]{shib2011-truncated_moment}):
\begin{equation}
  \label{eqn:se_kernel_expansion}
  R(\varepsilon,\omega) \approx R_0(\varepsilon) + \omega R_1(\varepsilon),
\end{equation}
which allows us to write Eqns.~\ref{eqn:se_factor} and \ref{eqn:se_opacity} as
\begin{align}
  \label{eqn:se_factor_expansion}
  \Phi_{{\rm SE},i} &=  \frac{(2\pi)^3}{\varepsilon^3}\frac{1}{4\pi}
  J(\varepsilon) -
  \frac{1}{\mathscr{K}_{{\rm SE},i}} \ell^\alpha H_\alpha(\varepsilon)\\
  \label{eqn:se_opacity_expansion}
  \mathscr{K}_{{\rm SE},i} &= 4\pi \frac{\varepsilon^3}{(2\pi)^3}
  \big(R_0(\varepsilon)+R_1(\varepsilon)\big).
\end{align}

\subsection{Inelastic scattering}
\label{ssec:sources_si}
(e.g. \cite[Eqn.~A6]{brue1985-core_collapse}).
